---
title: "Statistical Inference Course Project"
author: "Chandan Mahodi"
date: "5 January 2019"
output:
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Part1
##1. Show the sample mean and compare it to the theoretical mean of the distribution.

```{r part1_1}
lambda<-0.2
mysim<-matrix(rexp(1000*40,lambda),nrow=1000,ncol=40)
```

###The average of the 1000 means of the 40 exponentials :
```{r Sample_Mean}
mymean<-apply(mysim,1,mean)
mean(mymean)
```

###The Theoretical mean:
```{r Theoretical_Mean}
1/lambda
```
###Comparing Theoretical and sample mean:
````{r Hisotgram_comparison_mean}
hist(mymean,main="Distribution of 1000 means\n of 40 exponentials",xlab="Means",ylab="Frequency Of Means",breaks = 20)
abline(v=1/lambda,col="green",lwd=5)
```

We could see that both the center of the means distribution and the green vertical line of theoretical mean
are very closed to each other.

##2. Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.

###The average of the 1000 variances of the 40 exponentials :
```{r Sample_Variance}
myvar<-apply(mysim,1,var)
mean(myvar)
```

###The Theoretical var:
```{r Theoretical_Variance}
TheVar<-(1/lambda)^2
TheVar
```
###Comparing Theoretical and sample Variance:
````{r Hisotgram_comparison_var}
hist(myvar,main="Sample distribution of 1000 variances\n of 40 exponentials",xlab="Variances",ylab="Frequency Of Variances",breaks = 20)
abline(v=TheVar,col="green",lwd=5)
```

We could see that both the center of the sample Variances distribution and the green vertical line of theoretical Variance are very closed to each other.

##3. Show that the distribution is approximately normal.

```{r Normal_Distribution_Proof}
par(mfrow=c(2,1))
hist(mymean,breaks = 20,main="Distribution of Means of exponentials",xlab="Means",ylab="Frequency")
hist(rnorm(1000,mean=1/lambda,sd=1/lambda),breaks = 20,main="Normal Distribution",xlab="Normal Variables",ylab="Frequency")
```

The above graphs show how the distribution of means of exponentials is approximately a normal distribution with the same Mean and same Standard deviation.


